Mirror, mirror on the wall. Who's the best match of them all?

### 2023-11-22 PNAS/index.Rnw
added citation, partial matches

\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
\usepackage{cleveref}
\usepackage[dvipsnames]{xcolor} % colors
\newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
\newcommand{\hh}[1]{{\textcolor{ForestGreen}{#1}}}
\templatetype{pnasbriefreport} % Choose template
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

@@

This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{@fine2006}: one of the issues leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.

@@

The multiple comparison problems in forensics are distinct from the familywise error rate and false discovery rate problems that have been previously addressed in major ways: our tests are sequential, not concurrent, with dependencies depending on both the testing procedure and the data. In addition, an identification might be sufficient to end the search; as a result, this problem is more akin to a ``search until you find" problem than a situation where a family-wise error rate or false discovery rate is applicable - we are interested in the probability of at least one error in a series of sequential comparisons.

The multiple testing problem is closely related to issues which have caused embarrassing false accusations and mistrust of forensic pattern evidence: partial matches and database searches. As we assemble research databases for explicitly making many comparisons and tackle problems which implicitly involve multiple comparisons, we must confront these issues before more embarrassing missteps lead the public to question the validity of forensic science.

In this paper, we consider the multiple comparisons problem that arises from a relatively simple toolmark comparison: matching a cut wire to a wire-cutting tool. We describe the approach and calculate the number of comparisons performed, and then assess the likelihood of a false positive error considering the multiple comparisons performed, using error rates derived from published black-box studies.

>>>>

\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
\usepackage{cleveref}
\usepackage[dvipsnames]{xcolor} % colors
\newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
\newcommand{\hh}[1]{{\textcolor{ForestGreen}{#1}}}
\templatetype{pnasbriefreport}
%\templatetype{pnasbriefreport} % Choose template
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

@@

This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehlerFingerprintErrorRate2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
%
\hh{In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.}
%
\hh{In order to avoid embarassing false accusation and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.}
%
\hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria. }
\hh{Another source of (implicit) multiple comparisons are partial matches. }

\hh{XXX director's notes: XXX what is different between full and partial matches is that error rates in the literature are generally reported based on full length comparisons, where macro features are used to anchor comparisons (align along grooves, then zoom in and only 'wiggle' a bit to the left and right to adjust at the micro-scale). These kind of adjustments are 'baked' into published error rates. }

@@

%The multiple comparison problems in forensics are distinct from the familywise error rate and false discovery rate problems that have been previously addressed in major ways: our tests are sequential, not concurrent, with dependencies depending on both the testing procedure and the data. In addition, an identification might be sufficient to end the search; as a result, this problem is more akin to a ``search until you find" problem than a situation where a family-wise error rate or false discovery rate is applicable - we are interested in the probability of at least one error in a series of sequential comparisons.

%The multiple testing problem is closely related to issues which have caused embarrassing false accusations and mistrust of forensic pattern evidence: partial matches and database searches. %As we assemble research databases for explicitly making many comparisons and tackle problems which implicitly involve multiple comparisons, we must confront these issues before more embarrassing missteps lead the public to question the validity of forensic science.

Here, we consider the multiple comparisons problem that arises from a relatively simple toolmark comparison: matching a cut wire to a wire-cutting tool. We describe the approach and calculate the number of comparisons performed, and then assess the likelihood of a false positive error considering the multiple comparisons performed, using error rates derived from published black-box studies.
